DNAm Hierarchical Model - Prior Sensitivity Analysis 

### 2021-02-19 Paper_Figures.Rmd
added boxplots

# plot the two conservations score prob sums
genemask = !(FullAnnotation$UCSC_RefGene_Name=="")
p_scores <- post_histogram(CpGscore_full$median,genemask,0.2,"CpG Conservation score") +
  theme(aspect.ratio=1/4,legend.position = "none") +
  scale_x_continuous(limits=c(-8,8),breaks=c(-8,-4,0,4,8)) + 
  scale_y_continuous(breaks=c(0,0.1,0.2,0.3,0.4,0.5))
print(p_scores)
# ggsave("CpGscores_overlay.svg")

# Fig 4B: plot gene scores and adjusted bootstrap distribution
p1 <- permed_data %>% ggplot(aes(x=score_median)) + geom_histogram(aes(y=..count..), binwidth=0.1, fill="lightgray",color = "grey50") + xlim(c(-3,7)) + theme_classic() + ylab("Number of genes") + theme(aspect.ratio = 1)
p2 <- permed_data %>% ggplot(aes(x=score_median_v3)) + geom_histogram(aes(y=..count..), binwidth=0.02, fill="lightgray", color="grey50") + xlim(c(0,1)) + theme_classic() + ylab("Number of genes") + theme(aspect.ratio = 1)
p_genes <- ggarrange(p1,p2,nrows=2,ncol=1)
# ggsave("gene_scores_bootstrap.svg")
```

@@

# Fig 3B: plot CpG island relation score distributions
prepped_data %>% select(Name, Relation_to_UCSC_CpG_Island, score_median) %>%
  dplyr::rename(ir = Relation_to_UCSC_CpG_Island, score=score_median) %>%
  mutate(ir = case_when(is.na(ir) ~ "Sea",
                        str_detect(ir, "^S") ~ str_replace(ir, "^S_", ""),
                        str_detect(ir, "^N") ~ str_replace(ir, "^N_", ""),
                        TRUE ~ ir)) %>%
  mutate(ir = factor(ir, levels = c("Island", "Shore", "Shelf", "Sea"))) %>%
  #mutate(ir = factor(ir, levels = c("Sea", "Shelf", "Shore", "Island"))) %>%
  ggplot() + geom_violin(aes(x=ir,y=score,fill=ir),draw_quantiles = 0.5) + theme_classic() + scale_y_continuous(limits=c(-5,5),breaks=c(-4,-2,0,2,4)) + theme(aspect.ratio=1/4,legend.position = "none")
# ggsave("CpGscore_islands_violin.svg")
```

@@

# Fig 3C: plot functional region score distributions
temp <- prepped_data %>% select(Name, UCSC_RefGene_Group, score_median) %>%
  separate_rows(UCSC_RefGene_Group, sep=";") %>%
  distinct() %>%
  dplyr::rename(fr = UCSC_RefGene_Group, score=score_median) %>%
  mutate(fr = case_when(fr=="" ~ "NA",
                        str_detect(fr, "5URT") ~ "5'UTR",
                        str_detect(fr, "3UTR") ~ "3'UTR",
                        TRUE ~ fr)) %>%
  distinct()
#keepregions = c("TSS1500","Body", "3'UTR","1stExon","TSS200","5'UTR","ExonBnd")
keepregions = c("TSS200","1stExon", "5'UTR","TSS1500","Body","3'UTR","ExonBnd")
temp %>%
  mutate(fr = factor(fr, levels = keepregions)) %>%
  filter(fr %in% keepregions) %>%
  ggplot() + geom_violin(aes(x=fr,y=score,fill=fr), draw_quantiles = 0.5) + theme_classic() + scale_y_continuous(limits=c(-5,5),breaks=c(-4,-2,0,2,4,6)) + theme(aspect.ratio=1/4,legend.position = "none") + geom_hline(yintercept = 0, color="black", alpha=0.2)
# ggsave("CpGscore_regions_violin.svg")
#rm(temp)

@@

# plot violins of: all genes, COSMIC, AGO
perm_temp <- permed_data %>% select(c(UCSC_RefGene_Name, score_median,score_median_v3,n))
genes_dbs <- rbind(perm_temp %>% mutate(group="all"),
              perm_temp %>% filter(UCSC_RefGene_Name %in% (cancergenes1 %>% pull(`Gene Symbol`))) %>% mutate(group="COSMIC"),
              perm_temp %>% filter(UCSC_RefGene_Name %in% (cancergenes2 %>% pull(CancerGenes))) %>% mutate(group="AGO")
              )
# Fig 4B: plot violins
pDB <- genes_dbs %>% mutate(group = factor(group, levels = c("all", "AGO", "COSMIC"))) %>%
  ggplot() + geom_violin(aes(x=score_median,y=group,fill=group), draw_quantiles = 0.5) + scale_x_continuous(limits=c(-2,4),breaks = c(-2,0,2,4)) + theme_classic() + ylab("Database") + xlab("Gene Conservation Score") + geom_vline(xintercept=0, color="black", alpha=0.2) + theme(aspect.ratio=60/156.3,legend.position = "none")
# add hashmarks for each gene (doesn't look great)
marks <- genes_dbs %>% mutate(x=score_median, y=if_else(group=="COSMIC",3,if_else(group=="AGO",2,1))) %>% select(x,y)
marks2 <- genes_dbs %>% filter(score_median_v3 > 0.95) %>% mutate(x=score_median, y=if_else(group=="COSMIC",3,if_else(group=="AGO",2,1))) %>% select(x,y)
pDB + geom_segment(data=marks2, aes(x=x, xend=x, y=y, yend=y+0.2), color="red") + geom_segment(data=marks, aes(x=x, xend=x, y=y, yend=y+0.1), color="black", alpha=0.5)
# ggsave("database_violins.svg")

>>>>

# plot the two conservations score prob sums
genemask = !(FullAnnotation$UCSC_RefGene_Name=="")
p_scores <- post_histogram(CpGscore_full$median,genemask,0.2,"CpG Conservation score") +
  theme(aspect.ratio=1/4,legend.position = "none") +
  scale_x_continuous(limits=c(-8,8),breaks=c(-8,-4,0,4,8)) + 
  scale_y_continuous(breaks=c(0,0.1,0.2,0.3,0.4,0.5))
print(p_scores)

df = data.frame(med=CpGscore_full$median, mask=genemask)
p_scores <- ggplot(data=df) + geom_boxplot(aes(x=med, y = genemask, fill = genemask), outlier.shape = NA) +
  scale_x_continuous(limits=c(-8,8),breaks=c(-8,-4,0,4,8)) + xlab("CpG score") + ylab("Gene-associated") + theme_classic() + theme(aspect.ratio=1/4,legend.position = "none")
print(p_scores)
# ggsave("CpGscores_overlay.svg")

# Fig 4B: plot gene scores and adjusted bootstrap distribution
p1 <- permed_data %>% ggplot(aes(x=score_median)) + geom_histogram(aes(y=..count..), binwidth=0.1, fill="lightgray",color = "grey50") + xlim(c(-3,7)) + theme_classic() + ylab("Number of genes") + theme(aspect.ratio = 1)
p2 <- permed_data %>% ggplot(aes(x=score_median_v3)) + geom_histogram(aes(y=..count..), binwidth=0.02, fill="lightgray", color="grey50") + xlim(c(0,1)) + theme_classic() + ylab("Number of genes") + theme(aspect.ratio = 1)
p_genes <- ggarrange(p1,p2,nrows=1,ncol=2)
# ggsave("gene_scores_bootstrap.svg")
```

@@

# Fig 3B: plot CpG island relation score distributions
prepped_data %>% select(Name, Relation_to_UCSC_CpG_Island, score_median) %>%
  dplyr::rename(ir = Relation_to_UCSC_CpG_Island, score=score_median) %>%
  mutate(ir = case_when(is.na(ir) ~ "Sea",
                        str_detect(ir, "^S") ~ str_replace(ir, "^S_", ""),
                        str_detect(ir, "^N") ~ str_replace(ir, "^N_", ""),
                        TRUE ~ ir)) %>%
  #mutate(ir = factor(ir, levels = c("Island", "Shore", "Shelf", "Sea"))) %>%
  #ggplot() + geom_violin(aes(x=ir,y=score,fill=ir),draw_quantiles = 0.5) + theme_classic() + 
  mutate(ir = factor(ir, levels = c("Sea", "Shelf", "Shore", "Island"))) %>%
  ggplot() + geom_boxplot(aes(y=ir,x=score, fill=ir), outlier.shape = NA) + theme_classic() +
  scale_x_continuous(limits=c(-5,5),breaks=c(-4,-2,0,2,4)) + theme(aspect.ratio=1/4,legend.position = "none")
# ggsave("CpGscore_islands_violin.svg")
ggsave("CpGscore_islands_boxplot.svg")
```

@@

# Fig 3C: plot functional region score distributions
temp <- prepped_data %>% select(Name, UCSC_RefGene_Group, score_median) %>%
  separate_rows(UCSC_RefGene_Group, sep=";") %>%
  distinct() %>%
  dplyr::rename(fr = UCSC_RefGene_Group, score=score_median) %>%
  mutate(fr = case_when(fr=="" ~ "NA",
                        str_detect(fr, "5URT") ~ "5'UTR",
                        str_detect(fr, "3UTR") ~ "3'UTR",
                        TRUE ~ fr)) %>%
  distinct()
#regions = c("TSS1500","Body", "3'UTR","1stExon","TSS200","5'UTR","ExonBnd")
regions = rev(c("TSS200","1stExon", "5'UTR","TSS1500","Body","3'UTR","ExonBnd"))
temp %>%
  mutate(fr = factor(fr, levels = regions)) %>%
  filter(fr %in% regions) %>%
  #ggplot() + geom_violin(aes(x=fr,y=score,fill=fr), draw_quantiles = 0.5) + theme_classic() +
  ggplot() + geom_boxplot(aes(y=fr,x=score, fill=fr), outlier.shape = NA) + theme_classic() +
  scale_x_continuous(limits=c(-5,5),breaks=c(-4,-2,0,2,4)) + theme(aspect.ratio=1/4,legend.position = "none") + geom_vline(xintercept = 0, color="black", alpha=0.2)
# ggsave("CpGscore_regions_violin.svg")
ggsave("CpGscore_regions_boxplot.svg")
#rm(temp)3

@@

# plot violins of: all genes, COSMIC, AGO
perm_temp <- permed_data %>% select(c(UCSC_RefGene_Name, score_median,score_median_v3,n))
genes_dbs <- rbind(perm_temp %>% mutate(group="all"),
              perm_temp %>% filter(UCSC_RefGene_Name %in% (cancergenes1 %>% pull(`Gene Symbol`))) %>% mutate(group="COSMIC"),
              perm_temp %>% filter(UCSC_RefGene_Name %in% (cancergenes2 %>% pull(CancerGenes))) %>% mutate(group="AGO")
              )
# Fig 4C: plot distribution
pDB <- genes_dbs %>% mutate(group = factor(group, levels = c("all", "AGO", "COSMIC"))) %>%
  #ggplot() + geom_violin(aes(x=score_median,y=group,fill=group), draw_quantiles = 0.5) +
  ggplot() + geom_boxplot(aes(x=score_median,y=group,fill=group), outlier.shape = NA) +
  scale_x_continuous(limits=c(-2,4),breaks = c(-2,0,2,4)) + theme_classic() + ylab("Database") + xlab("Gene Conservation Score") + geom_vline(xintercept=0, color="black", alpha=0.2) + theme(aspect.ratio=60/156.3,legend.position = "none")
# # add hashmarks for each gene (doesn't look great)
# marks <- genes_dbs %>% mutate(x=score_median, y=if_else(group=="COSMIC",3,if_else(group=="AGO",2,1))) %>% select(x,y)
# marks2 <- genes_dbs %>% filter(score_median_v3 > 0.95) %>% mutate(x=score_median, y=if_else(group=="COSMIC",3,if_else(group=="AGO",2,1))) %>% select(x,y)
# pDB + geom_segment(data=marks2, aes(x=x, xend=x, y=y, yend=y+0.2), color="red") + geom_segment(data=marks, aes(x=x, xend=x, y=y, yend=y+0.1), color="black", alpha=0.5)
pDB
# ggsave("database_violins.svg")
ggsave("database_boxplot.svg")

### 2021-02-19 prior_sensitivity.Rmd
added boxplots

>>>>

```{r posterior_sensitivity, echo=FALSE, comment=NA}

stanFile = "FullResultsTCGA_relaxgamma3.Rdata"
load(here(stanFile))

mu_ref <- mu_full
nu_ref <- nu_full
sigmaP_ref <- sigmaP_full
sigmaPT_ref <- sigmaPT_full
sigmaT_ref <- sigmaT_full
sigmaE_ref <- sigmaE_full

for (i in c(1:4)) {
  stanFile = switch(i,
                    "FullResultsTCGA_gamma.Rdata",
                    "FullResultsTCGA_relaxgamma2.Rdata",
                    "FullResultsTCGA_relaxgamma3.Rdata",
                    "FullResultsTCGA_relaxgamma5.Rdata")
  print(stanFile)
  load(here(stanFile))
  print("loaded")

  for (j in c(1:6)) {
    # choose parameter (first get reference dist (3x), then prior relaxed dist)
    breaks <- switch(j,
                     seq(-5,5,0.1),
                     seq(-6,6,0.12), #5.5,.11
                     seq(0,3,0.03), #2.6
                     seq(0,3,0.03),
                     seq(0,3,0.03), #1.2
                     seq(0,3,0.03)) #2.5

    refdata <- switch(j,
                   mu_ref$p50,
                   nu_ref$p50,
                   sigmaP_ref$p50,
                   sigmaPT_ref$p50,
                   sigmaT_ref$p50,
                   sigmaE_ref$p50)
    Q <- hist(refdata, plot=FALSE, breaks=breaks) # run hist to get density and midpoints
    x <- Q$mids

    data <- switch(j,
                   mu_full$p50,
                   nu_full$p50,
                   sigmaP_full$p50,
                   sigmaPT_full$p50,
                   sigmaT_full$p50,
                   sigmaE_full$p50)
    P <- hist(data, plot=FALSE, breaks=breaks) # run hist to get density and midpoints

    Pi <- P$density[which(P$density>0 & Q$density>0)]
    Qi <- Q$density[which(P$density>0 & Q$density>0)]
    KL_temp[j] = sum(Pi * log(Pi/Qi))

    # BCD
    dx <- x[2]-x[1]
    BC_temp[j] = -log(sum(sqrt(P$density*Q$density)*dx))

    # EMD
    A = cbind(P$density,P$mids)
    B = cbind(Q$density,Q$mids)
    EM_temp[j] = emd(A,B)

print("N_eff")
kable(n_eff_mat, digits=2)
print("Rhat")
kable(Rhat_mat, digits=3)
print("KLdiv");kable(KLdiv_mat, digits=3);print("BCdist");kable(BCdist_mat, digits=3);print("EMdist"); kable(EMdist_mat, digits=3)
